The Implications of Local Fluctuations in the Galactic Midplane for Dynamical Analysis in the \textit{Gaia} Era

### 2019-07-26 paper/ms.tex
Add some loose references

\textbf{The true Galactic potential is not strictly axisymmetric, beyond even
the typically-quoted gas clouds, spiral arms, and bar(s). The presence of disk
oscillations has been known
%mm for some time, observed in 1957 for
% H~% \textsc{i} in the outer disk \citep{1957AJ.....62...93K}.
    since its observation in H~{sc i} by \citet{1957AJ.....62...93K}.
The presence of a North-South
asymmetry in stellar number density and velocity in the solar neighborhood has
been found in
% mm [should these multiple acronyms be expanded?]
SDSS data \citep{2012ApJ...750L..41W}, and in velocity in the
LAMOST survey \citep{2013ApJ...777L...5C}, and RAVE data
\citep{2013MNRAS.436..101W}. Beyond the solar neighborhood, low-latitude
overdensities have been observed, e.g., the Monoceros Ring
\citep{2002ApJ...569..245N, 2003MNRAS.340L..21I, 2003ApJ...594L.119C,
2014ApJ...791....9S, 2016ApJ...825..140M}, an overdensity in the direction of
the Triangulum and Andromeda galaxies \citep{2007ApJ...668L.123M,
2014ApJ...793...62S, 2015ApJ...801..105X, 2015MNRAS.452..676P} and A13
\citep{2010ApJ...722..750S, 2017ApJ...844...74L}, and in the direction of the
Galactic Center \citep{2014Natur.509..342F}. There is evidence that these
stellar populations originated from the disk \citep{2018ApJ...854...47S,
2018Natur.555..334B}. See, e.g., \citet{2018MNRAS.481..286L} for a discussion
of these observations.}

\textbf{With the vast improvement in the quality of phase-space measurements
due to \textit{Gaia} the assumption of axisymmetry is increasingly inadequate
\citep[e.g.,][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. The presence of the
phase-space spiral itself is evidence of disk oscillations
\citep{2018Natur.561..360A}, but the fact that its strength changes across the
disk is further evidence \citep{2019MNRAS.486.1167B}. Even if the axisymmetric
assumption were close enough for many purposes, the parameters used in
axisymmetric models of the Galactic potential may be inadequately constrained
by current observations.}

>>>>

\textbf{The true Galactic potential is not strictly axisymmetric, beyond even
the typically-quoted gas clouds, spiral arms, and bar(s). The presence of disk
oscillations has been known
%mm for some time, observed in 1957 for
% H~% \textsc{i} in the outer disk \citep{1957AJ.....62...93K}.
    since its observation in H~{sc i} by \citet{1957AJ.....62...93K}.
The presence of a North-South
asymmetry in stellar number density and velocity in the solar neighborhood has
been found in
% mm [should these multiple acronyms be expanded?]
SDSS data \citep{2012ApJ...750L..41W}, and in velocity in the
LAMOST survey \citep{2013ApJ...777L...5C, 2015ApJ...801..105X}, and RAVE data
\citep{2013MNRAS.436..101W}. Beyond the solar neighborhood, low-latitude
overdensities have been observed, e.g., the Monoceros Ring
\citep{2002ApJ...569..245N, 2003MNRAS.340L..21I, 2003ApJ...594L.119C,
2014ApJ...791....9S, 2016ApJ...825..140M}, an overdensity in the direction of
the Triangulum and Andromeda galaxies \citep{2007ApJ...668L.123M,
2014ApJ...793...62S, 2015ApJ...801..105X, 2015MNRAS.452..676P} and A13
\citep{2010ApJ...722..750S, 2017ApJ...844...74L}, and in the direction of the
Galactic Center \citep{2014Natur.509..342F}. There is evidence that these
stellar populations originated from the disk \citep{2015MNRAS.452..676P, 2018ApJ...854...47S,
2018Natur.555..334B}. See, e.g., \citet{2018MNRAS.481..286L} for a discussion
of these observations.}

\textbf{With the vast improvement in the quality of phase-space measurements
due to \textit{Gaia} the assumption of axisymmetry is increasingly inadequate
\citep[e.g.,][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. The presence of the
phase-space spiral itself is evidence of disk oscillations
\citep{2018Natur.561..360A}, but the fact that its strength changes across the
disk is further evidence \citep{2019MNRAS.485.3134L, 2019MNRAS.486.1167B}. Even if the axisymmetric
assumption were close enough for many purposes, the parameters used in
axisymmetric models of the Galactic potential may be inadequately constrained
by current observations.}
